---
title: "Machine Learning Assignment"
author: "Bradley JK Smith"
date: "26 October 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

... goes here ...

## Get Data

This is already split into training and testing sets
Think about whether pre-processing is required
?? Need to fix any spelling errors in column names?? eg. #88 kurtosis\_picth\_dumbbell
Probably set a seed here 
(would only be required if there is some randomness is training eg. random forests)
```{r}
training <- read.csv('pml-training.csv')
testing <- read.csv('pml-testing.csv')

```

## Exploratory Analysis

- number of variables
- if/how to use user_name (eg. one user can follow the instructions better)
- if/how to use timestamp (eg. get better at following the instructions over time)

There are sensors on the arm, forearm, belt and dumbbell

How to deal with missing data (Is there missing data?)
Whether or not to create new features

Correlation between predictors - whether to use PCA


## Discussion of what type of model to use

Options are:

- regression
- trees
- xxx

Discuss the trade offs

accuracy vs interpretability

